---
title: "01-diff-exp-analysis"
output: html_document
date: "2023-12-12"
---

```{bash, engine.opts='-l'}
echo $PATH
```

# Bring tag-seq data in from gannet
```{bash}
mkdir ~/Github/byssus-exp-analysis/data/raw-data/
cd ~/Github/byssus-exp-analysis/data/raw-data/

wget -r \
--no-directories --no-parent \
-P . \
-A .fastq.gz https://gannet.fish.washington.edu/panopea/PSMFC-mytilus-byssus-pilot/20220405-tagseq/ \
--no-check-certificate
```

# Run fastqc on untrimmed files
```{bash}
mkdir ../output/fastqc/untrimmed/

/home/shared/FastQC-0.12.1/fastqc \
../data/raw-data/*.fastq.gz \
--outdir ../output/fastqc/untrimmed/ \
--quiet

```

# Run multiqc on untrimmed fastqc files
```{bash}

eval "$(/opt/anaconda/anaconda3/bin/conda shell.bash hook)"
conda activate

cd ~/Github/byssus-exp-analysis/output/fastqc/untrimmed/

multiqc .

```
# Download trimmed files
```{bash}
mkdir ../data/raw-trimmed/
wget -e robots=off -np --input-file=../data/download.txt -P ../data/raw-trimmed/
```


# Run fastqc on trimmed files
```{bash}
mkdir ../output/fastqc/trimmed/
/home/shared/FastQC-0.12.1/fastqc \
../data/raw-trimmed/*.fastq.gz \
--outdir ../output/fastqc/trimmed/ \
--quiet

```

# Run multiqc on trimmed fastqc files
```{bash}

eval "$(/opt/anaconda/anaconda3/bin/conda shell.bash hook)"
conda activate

cd ../output/fastqc/trimmed/

multiqc .

```

# get transcriptome

```{r, engine='bash'}
cd ../data
curl -O https://owl.fish.washington.edu/halfshell/genomic-databank/Mtros-hq_transcripts.fasta
```


# Create indices for transcriptome with Kallisto
```{bash}
# Build index

# Execute the Kallisto tool located at the specified path
/home/shared/kallisto_linux-v0.50.1/kallisto \
index \
-t 20 \
-i ../output/Mtros-hq_transcripts.index \
../data/Mtros-hq_transcripts.fasta
```

# Quantify the indices
```{bash}
# Set the paths
DATA_DIRECTORY="../data/raw-trimmed/"
KALLISTO_INDEX="../output/Mtros-hq_transcripts.index"
OUTPUT_DIRECTORY="../output/kallisto-quant"


# Iterate over all .fq.gz files in the data directory
for FILE in "$DATA_DIRECTORY"/*.fastq.gz; do
    # Extract the base name of the file for naming the output folder
    BASENAME=$(basename "$FILE" _L099_R1_cmb.trim.fastq.gz)

    # Create output directory for this sample
    SAMPLE_OUTPUT="$OUTPUT_DIRECTORY/$BASENAME"
    mkdir -p "$SAMPLE_OUTPUT"

    # Run Kallisto quantification
    /home/shared/kallisto_linux-v0.50.1/kallisto quant -i "$KALLISTO_INDEX" -o "$SAMPLE_OUTPUT" \
        --single -t 20 -l 65 -s 2 "$FILE"
done

echo "Kallisto quantification complete."
```
# Generate count matrices
```{bash}
     
perl /home/shared/trinityrnaseq-v2.12.0/util/abundance_estimates_to_matrix.pl \
--est_method kallisto \
--gene_trans_map 'none' \
--out_prefix ../output/kallisto-quant/abundance-matrices \
--name_sample_by_basedir \
../output/kallisto-quant/*/abundance.tsv 

```
# View abundance matrix
```{bash}
head ../output/kallisto-quant/abundance-matrices.isoform.counts.matrix
```
# Make counts matrix
```{r, eval = TRUE}
countmatrix <- read.delim("../output/kallisto-quant/abundance-matrices.isoform.counts.matrix", header = TRUE, sep = '\t')
rownames(countmatrix) <- countmatrix$X
countmatrix <- countmatrix[,-1]
head(countmatrix)
```

# Make count matrix into integers
```{r, eval=TRUE}
countmatrix <- round(countmatrix, 0)
str(countmatrix)
```




