se---
title: "Getting_GOSlims"
output: html_document
date: "2024-04-03"
---
```{r}
library(tidyverse)

GOSlims <- read.csv("https://gannet.fish.washington.edu/spartina/project-oyster-oa/Haws/GO-annotation/GO-GOslim.sorted", sep="\t", header = F)

colnames(GOSlims) <- c("GO.IDs", "Pathway", "Process", "Letter")
```

#FDO
```{r}

FDO_sigs_ID_cleaned_separated <- separate_rows(FDO_sigs_ID, GO.IDs, sep = ";")
FDO_sigs_ID_cleaned_separated$GO.IDs <- trimws(FDO_sigs_ID_cleaned_separated$GO.IDs)


```

### Map Biological Process GOslims to GO IDs
```{r BP-GOslims-to-GOIDs}
# Vector of GO IDs
go_ids <- FDO_sigs_ID_cleaned_separated$V1

# Create custom collection with our GO IDs
myCollection <- GOCollection(go_ids)

# Pull out GOslims from GOslim obo file
slim <- getOBOCollection(fl)

# Get Biological Process (BP) GOslims for our GO IDs
slimdf <- goSlim(myCollection, slim, "BP", verbose)

# Convert to list
gomap <- as.list(GOBPOFFSPRING[rownames(slimdf)])

# Map our GOIDs to GOslims
mapped <- lapply(gomap, intersect, ids(myCollection))

# sapply needed to apply paste() to create semi-colon delimited values
slimdf$ids <- sapply(lapply(gomap, intersect, ids(myCollection)), paste, collapse=";")

# Remove "charcter(0) string from "ids" column
slimdf$ids[slimdf$ids == "character(0)"] <- ""

# Add self-matching GOIDs to "ids" column, if not present
updated_slimdf <- slimdf

for (go_id in go_ids) {
  # Check if the go_id is present in the row names
  if (go_id %in% rownames(updated_slimdf)) {
    # Check if the go_id is not present in the ids column
    # Also removes white space "trimws()" and converts all to upper case to handle
    # any weird, "invisible" formatting issues.
    if (!go_id %in% trimws(toupper(strsplit(updated_slimdf[go_id, "ids"], ";")[[1]]))) {
      # Append the go_id to the ids column with a semi-colon separator
      if (length(updated_slimdf$ids) > 0 && nchar(updated_slimdf$ids[nrow(updated_slimdf)]) > 0) {
        updated_slimdf[go_id, "ids"] <- paste0(updated_slimdf[go_id, "ids"], "; ", go_id)
      } else {
        updated_slimdf[go_id, "ids"] <- go_id
      }
    }
  }
}

str(updated_slimdf)
```

### Create GSEAbase GeneSetCollection

```{r gene-set-collection}

# Group by unique GO ID
grouped_df <- my_data_cleaned_separated %>% group_by(V1)%>% summarise(V2 = paste(V2, collapse = ","))

# convert the data frame to a named list
my_list <- as.list(grouped_df$V2)
names(my_list) <- grouped_df$V1

gsc <- GeneSetCollection(mapply(function(geneIds, GOID) {
         GeneSet(geneIds, collectionType=GOCollection(GOID), setName=GOID)
       }, my_list, names(my_list)))

gsc
```

### Map genes to GOslims
```{r map-genes-to-GOslims}
mapped_df <- mappedIds(updated_slimdf, myCollection, GOBPOFFSPRING, gsc)

# Provide column name for first column
mapped_df <- cbind(GOslim.BP = rownames(mapped_df), mapped_df)
rownames(mapped_df) <- NULL

# Rename "ids" column
mapped_df <- mapped_df %>% 
  dplyr::rename(GO.IDs = ids)

mapped_df

str(mapped_df)
```

### Flatten GOslims to have one row per gene

Will make future joins to other data sets easier.

Done in two steps for easier readability:

1. Flattened just by gene and retains all info.

2. Flattened by gene with GOslims combined per gene. Drops GO IDs.


For the second version:

We use n() to count the number of rows within each group.

We select the first Term value within each group using Term[1].

.groups = 'drop' is used to remove the grouping information.

This code should create a dataframe with one row per gene, including the corresponding GOslim.BP entries in a single, semi-colon-delimited field, the first Term value within each group, and the count of rows in each group.

```{r flatten-GOslims-by-gene}
flattened_mapped_df <- mapped_df %>%
  dplyr::select(Genes, GOslim.BP, Term, GO.IDs) %>%
  separate_rows(Genes, sep = ",")
  

goslims_per_gene_df <- flattened_mapped_df %>%
  distinct(Genes, GOslim.BP, Term) %>%
  group_by(Genes) %>%
  summarise(GOslim.BP = paste(unique(GOslim.BP), collapse = ";"), 
            Term = paste(unique(Term), collapse = ";"),  # Concatenate Terms with semi-colon
            .groups = 'drop')
  

str(goslims_per_gene_df)
```

## Write mapped GOslims to file
```{r write-mapped-GOslims-to-file}
mapped_df %>%
  dplyr::select(GOslim.BP, Term, GO.IDs, Genes) %>% 
  write.table(file ="../output/17.1-GO-and-GOslim-CDS-Annotation/Cvir-CDS-GOslim.BP_term_GOIDs_genes.tab",
          row.names = FALSE,
          col.names = TRUE,
          quote = FALSE,
          sep = "\t"
          )
```

## Write flattened genes to file
```{r write-flattened-genes-to-file}
goslims_per_gene_df %>%
  write.table(file ="../output/17.1-GO-and-GOslim-CDS-Annotation/Cvir-CDS-GOslim.BP_per_gene.tab",
              row.names = FALSE,
              col.names = TRUE,
              quote = FALSE,
              sep = "\t"
              )
  
```