se---
title: "Getting_GOSlims"
output: html_document
date: "2024-04-03"
---
## Set GSEAbase location and download `goslim_generic.obo`
```{r download-goslim-obo}
library(GSEABase)
library(tidyverse)
goslims <- "goslim_generic.obo"
goslims_url <- "http://current.geneontology.org/ontology/subsets/goslim_generic.obo"
gseabase_location <- find.package("GSEABase")
goslim_obo_dest <- file.path(gseabase_location, "extdata", goslims, fsep = "/")
download.file(url = goslims_url,
              destfile = goslim_obo_dest)

fl <- system.file("extdata", goslims, package="GSEABase")
```

# Create functions
```{r create-functions}

# Function for mapping GOIDs to GOslims

mappedIds <- function(df, collection, OFFSPRING, goannotGSC) {
  # Split semi-colon delimited strings in df$ids
  df$ids_split <- strsplit(as.character(df$ids), ";")
  
  # Determine the maximum number of elements per row
  max_elements <- max(lengths(df$ids_split))
  
  # Fill in shorter rows with NA values
  df$ids_split <- lapply(df$ids_split, function(x) {
    length(x) <- max_elements
    x
  })
  
  # Combine the split strings into a matrix
  ids_mat <- do.call(rbind, df$ids_split)
  
  # Convert the matrix to a data frame
  ids_df <- as.data.frame(ids_mat, stringsAsFactors = FALSE)
  
  # Rename the columns of the data frame
  colnames(ids_df) <- paste0("ids_", 1:max_elements)
  
  # Combine the original data frame and the new data frame
  df <- cbind(df, ids_df)
  
  # Perform the matching operation
  mt <- match(toupper(trimws(df$ids_1)), toupper(trimws(names(goannotGSC))))
  
  # Add a new column to the data frame for the "Genes" result
  df$Genes <- NA_character_
  
  # Fill in the "Genes" column for rows that have a match
  df$Genes[!is.na(mt)] <- vapply(geneIds(goannotGSC)[mt[!is.na(mt)]], 
                                  paste, collapse = ";", character(1L))
  
   df <- df %>% dplyr::select(Count, Percent, Term, ids, Genes)
   
  # Return the modified data frame
  return(df)
}
```


```{r}

FDO_sigs_ID_cleaned_separated <- separate_rows(FDO_sigs_ID, GO.IDs, sep = ";")
FDO_sigs_ID_cleaned_separated$GO.IDs <- trimws(FDO_sigs_ID_cleaned_separated$GO.IDs)

# Vector of GO IDs
FDO_go_ids <- FDO_sigs_ID_cleaned_separated$GO.IDs

# Create custom collection with our GO IDs
FDOmyCollection <- GOCollection(FDO_go_ids)

# Pull out GOslims from GOslim obo file
slim <- getOBOCollection(fl)

# Get Biological Process (BP) GOslims for our GO IDs
slimdf <- goSlim(FDOmyCollection, slim, "BP", verbose)

# Convert to list
FDOgomap <- as.list(GOBPOFFSPRING[rownames(slimdf)])

# Map our GOIDs to GOslims
mapped <- lapply(FDOgomap, intersect, ids(FDOmyCollection))

# sapply needed to apply paste() to create semi-colon delimited values
slimdf$ids <- sapply(lapply(FDOgomap, intersect, ids(FDOmyCollection)), paste, collapse=";")

# Add self-matching GOIDs to "ids" column, if not present
updated_slimdf$ids <- rownames(slimdf)

for (go_id in go_ids) {
  # Check if the go_id is present in the row names
  if (go_id %in% rownames(updated_slimdf)) {
    # Check if the go_id is not present in the ids column
    # Also removes white space "trimws()" and converts all to upper case to handle
    # any weird, "invisible" formatting issues.
    if (!go_id %in% trimws(toupper(strsplit(updated_slimdf[go_id, "ids"], ";")[[1]]))) {
      # Append the go_id to the ids column with a semi-colon separator
      if (length(updated_slimdf$ids) > 0 && nchar(updated_slimdf$ids[nrow(updated_slimdf)]) > 0) {
        updated_slimdf[go_id, "ids"] <- paste0(updated_slimdf[go_id, "ids"], "; ", go_id)
      } else {
        updated_slimdf[go_id, "ids"] <- go_id
      }
    }
  }
}

str(updated_slimdf)
```

### Create GSEAbase GeneSetCollection

```{r gene-set-collection}

# Group by unique GO ID
grouped_df <- FDO_sigs_ID_cleaned_separated %>% group_by(GO.IDs) %>% summarise(geneID = paste(geneID, collapse = ","))

# convert the data frame to a named list
my_list <- as.list(grouped_df$geneID)
names(my_list) <- grouped_df$GO.IDs

gsc <- GeneSetCollection(mapply(function(geneIds, GOID) {
         GeneSet(geneIds, collectionType=GOCollection(GOID), setName=GOID)
       }, my_list, names(my_list)))

gsc
```

### Map genes to GOslims
```{r map-genes-to-GOslims}
mapped_df <- mappedIds(updated_slimdf, FDOmyCollection, GOBPOFFSPRING, gsc)

# Provide column name for first column
mapped_df <- cbind(GOslim.BP = rownames(mapped_df), mapped_df)
rownames(mapped_df) <- NULL

# Rename "ids" column
mapped_df <- mapped_df %>% 
  dplyr::rename(GO.IDs = ids)

mapped_df

str(mapped_df)
```

### Flatten GOslims to have one row per gene

Will make future joins to other data sets easier.

Done in two steps for easier readability:

1. Flattened just by gene and retains all info.

2. Flattened by gene with GOslims combined per gene. Drops GO IDs.


For the second version:

We use n() to count the number of rows within each group.

We select the first Term value within each group using Term[1].

.groups = 'drop' is used to remove the grouping information.

This code should create a dataframe with one row per gene, including the corresponding GOslim.BP entries in a single, semi-colon-delimited field, the first Term value within each group, and the count of rows in each group.

```{r flatten-GOslims-by-gene}
flattened_mapped_df <- mapped_df %>%
  dplyr::select(Genes, GOslim.BP, Term, GO.IDs) %>%
  separate_rows(Genes, sep = ",")
  

goslims_per_gene_df <- flattened_mapped_df %>%
  distinct(Genes, GOslim.BP, Term) %>%
  group_by(Genes) %>%
  summarise(GOslim.BP = paste(unique(GOslim.BP), collapse = ";"), 
            Term = paste(unique(Term), collapse = ";"),  # Concatenate Terms with semi-colon
            .groups = 'drop')
  

str(goslims_per_gene_df)
```

## Write mapped GOslims to file
```{r write-mapped-GOslims-to-file}
mapped_df %>%
  dplyr::select(GOslim.BP, Term, GO.IDs, Genes) %>% 
  write.table(file ="../output/17.1-GO-and-GOslim-CDS-Annotation/Cvir-CDS-GOslim.BP_term_GOIDs_genes.tab",
          row.names = FALSE,
          col.names = TRUE,
          quote = FALSE,
          sep = "\t"
          )
```

## Write flattened genes to file
```{r write-flattened-genes-to-file}
goslims_per_gene_df %>%
  write.table(file ="../output/17.1-GO-and-GOslim-CDS-Annotation/Cvir-CDS-GOslim.BP_per_gene.tab",
              row.names = FALSE,
              col.names = TRUE,
              quote = FALSE,
              sep = "\t"
              )
  
```


